\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,color,float}
\usepackage{graphicx,psfrag,epsf}
\usepackage{natbib}


\setlength{\oddsidemargin}{.15in} 
\setlength{\textwidth}{6.25in}
\setlength{\topmargin}{-0.25in}
\setlength{\headheight}{-0.15in}
\setlength{\textheight}{8.9in} 

\linespread{1.25}


\title{ST 705 Linear models and variance components \\ 
        Homework problem set 4}


\begin{document}
\maketitle

\begin{enumerate}

\item Let $V$ be a finite-dimensional inner product space over $\mathbb{C}$, and let $\{v_{1},\dots,v_{n}\}$ be an orthonormal basis for $V$.  
\begin{enumerate}
\item Show that for any $x,y \in V$,
\[
\langle x,y\rangle = \sum_{i=1}^{n}\langle x,v_{i} \rangle\overline{\langle y,v_{i} \rangle}.
\]
This is called Parseval's identity.
\item For $V = \mathbb{R}^{2}$ use Parseval's identity to prove the Pythagorean theorem.  Generalize this result to $\mathbb{R}^{n}$.
\end{enumerate}

\item Let $V$ be an inner product space over $\mathbb{C}$, and let $\{v_{1},\dots,v_{n}\} \subset V$ be orthonormal.
\begin{enumerate}
\item Prove that for any $x \in V$,
\[
\|x\|^{2} \ge \sum_{i=1}^{n} |\langle x,v_{i}\rangle|^{2}.
\]
This is called Bessel's inequality.
\item Show that Bessel's inequality is an equality if and only if $x \in \text{span}\{v_{1},\dots,v_{n}\}$.
\end{enumerate}  

\item The defining property of a projection matrix $A$ is that $A^{2} = A$ (recall the definition of the square of a matrix from your linear algebra course).  Establish the following facts.
\begin{enumerate}
\item If $A$ is a projection matrix, then all of its eigenvalues are either zero or one. 
\item If $A \in \mathbb{R}^{p\times p}$ is a projection and symmetric (i.e., an orthogonal projection matrix), then for every vector $v$ the projection $Av$ is orthogonal to $v - Av$.
\item $\text{tr}(A + B) = \text{tr}(A) + \text{tr}(B)$.
\item $\text{tr}(AB) = \text{tr}(BA)$.
\end{enumerate}

\item Show that a $p\times p$ matrix $A$ is symmetric and idempotent with rank $s$ if and only if there exists a $p\times s$ matrix $G$ with orthonormal columns such that $A = GG'$.  Note that $G$ is called a {\em semi-orthogonal} matrix.

\item Exercise 2.8 from Monahan.

\item Exercise 2.9 from Monahan.

\item Exercise 2.11 from Monahan.


\end{enumerate}






\end{document}