\documentclass[11pt]{article}
\usepackage{amsmath,amssymb,color,float}
\usepackage{graphicx,psfrag,epsf}
\usepackage{natbib}


\setlength{\oddsidemargin}{.15in} 
\setlength{\textwidth}{6.25in}
\setlength{\topmargin}{-0.25in}
\setlength{\headheight}{-0.15in}
\setlength{\textheight}{8.9in} 

\linespread{1.25}


\title{ST 705 Linear models and variance components \\ 
        Homework problem set 6}


\begin{document}
\maketitle

\begin{enumerate}

\item(2 points) Exercise 3.9 from Monahan.  Not necessary to do the ``(More practice)'' item. 

\item(2 points) Consider the model $Y_{ijk} = \mu + \alpha_{i} + \beta_{j} + \theta_{ij} + U_{ijk}$, with $k \in \{1,\dots,m_{ij}\}$, $i \in \{1,\dots,n\}$, $j \in \{1,\dots,m\}$, and $E(U_{ijk}) = 0$.  Find necessary and sufficient conditions for which $\lambda'\gamma$ is estimable for 
\[
\gamma =
\begin{pmatrix}
\mu & \alpha_{1} & \cdots & \alpha_{n} & \beta_{1} & \cdots & \beta_{m} & \theta_{11} & \cdots & \theta_{nm} \\
\end{pmatrix}'.
\]

\item(2 points) Prove the converse to Result 3.2 (recall that we saw two proofs of Result 3.2 in lecture).  That is, if the least squares estimator $\lambda'\widehat{\beta}$ is the same for all solutions $\widehat{\beta}$ to the normal equations, then $\lambda'\beta$ is estimable. 

\item(2 points) Let $V$ be a finite-dimensional inner product space over $\mathbb{C}$, and let $\{v_{1},\dots,v_{n}\}$ be an orthonormal basis for $V$.  
\begin{enumerate}
\item Show that for any $x,y \in V$,
\[
\langle x,y\rangle = \sum_{i=1}^{n}\langle x,v_{i} \rangle\overline{\langle y,v_{i} \rangle}.
\]
This is called Parseval's identity.
\item For $V = \mathbb{R}^{2}$ use Parseval's identity to prove the Pythagorean theorem.  Generalize this result to $\mathbb{R}^{n}$.
\end{enumerate}

\item(2 points) Let $V$ be an inner product space over $\mathbb{C}$, and let $\{v_{1},\dots,v_{n}\} \subset V$ be orthonormal.
\begin{enumerate}
\item Prove that for any $x \in V$,
\[
\|x\|^{2} \ge \sum_{i=1}^{n} |\langle x,v_{i}\rangle|^{2}.
\]
This is called Bessel's inequality.
\item Show that Bessel's inequality is an equality if and only if $x \in \text{span}\{v_{1},\dots,v_{n}\}$.
\end{enumerate}  

\end{enumerate}






\end{document}